---
output:
  xaringan::moon_reader:
    self_contained: false
    css: ["default", "modern.css"]
    nature:
      highlightStyle: github
      highlightLines: true
      ratio: "16:9"
      transition: "fade"
includes:
  in_header: fonts.html
---
class: center, middle, title-slide

# Unaligned Decoupling  
### *How U.S. Allies Responded to Washington’s Trade War with China*
#### Ikhyun Lee  
#### School of International Service, American University
#### April 9, 2025

---
# Overview

**Background**  
	•	In 2018, the United States launched a trade war with China, aiming to economically decouple
<br>
	•	The U.S. has consistently tried to build a coalition of like-minded partners to reduce dependence on China

***

**Research Question**  
Did U.S. allies align with U.S. strategy and reduce their own strategic imports from China?

***

**Key Finding**  
The U.S. reduced strategic imports after 2018.  
Most allies? They went the opposite direction.

---
# Data

**Data Source**: UN Comtrade (cleaned and processed in R)
***
**Scope of Analysis**
- **Years**: 2014–2022
- **Countries**: United States, Japan, South Korea, United Kingdom, Germany, France, Australia
- **Products**: 13 strategic HS codes (e.g., semiconductors, aircraft parts, telecom equipment)
***
**Unit of Analysis**: Country–year–product imports from China (n = 787)

---
<h2 style="font-size: 50px; font-weight: 700; margin-bottom: 10px;">
Analysis 1: Total Strategic Imports from China
</h2>

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=4.4, out.width='100%', out.extra='style="margin-bottom:0;"'}
library(ggplot2)
library(scales)
library(dplyr)
library(stargazer)

df <- read.csv("cleaned_trade_data.csv")
df_summary <- df %>%
  group_by(year, importer) %>%
  summarise(total_import = sum(trade_value, na.rm = TRUE), .groups = "drop")

# Plot total imports by year-country
ggplot(df_summary, aes(x = year, y = total_import / 1e9, color = importer)) +
  geom_line(size = 1) +
  geom_point() +
  geom_vline(xintercept = 2018, linetype = "dashed", color = "red") +
  scale_y_continuous(labels = label_number(suffix = "")) +
  labs(
    title = "Total Strategic Goods Imports from China",
    subtitle = "2014–2022 (Dashed Line: U.S.-China Trade War Begins)",
    x = "Year", y = "Import Value (USD, Billions)", color = "Country"
  ) +
  theme_minimal()
```

---
# Analysis 2: Post-2018 Divergence (Aggregated Interaction Model)

## *What does this model estimate?*

- Whether U.S. allies **diverged from** the U.S. after the trade war began.
- DV: Log of strategic imports from China(USD)
- IV: `Post-2018 × Country` interactions
- Benchmarks each country **relative to the U.S.**
 
<br>
🟢 Positive coefficients = divergence from U.S. post-2018
<br>
🔴 Negative coefficients = alignment with U.S. decoupling

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(broom)
library(knitr)
library(kableExtra)

# Use aggregated country-year data
df_agg <- df %>%
  group_by(importer, year) %>%
  summarise(
    trade_value = sum(trade_value, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  filter(trade_value > 0) %>%
  mutate(
    log_trade = log(trade_value),
    post2018 = ifelse(year >= 2018, 1, 0),
    importer = relevel(factor(importer), ref = "United States")
  )

model_agg <- lm(log_trade ~ post2018 * importer, data = df_agg)
```
---
# Results
<div style="display: flex; justify-content: space-between; align-items: flex-start; gap: 40px;">

<!-- LEFT: Table -->
<div style="flex: 1.2;">
```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}

# Tidy + add significance stars
table_df <- tidy(model_agg) %>%
  mutate(term = case_when(
    term == "post2018" ~ "USA (Post-2018)",
    term == "importerAustralia" ~ "Australia",
    term == "importerFrance" ~ "France",
    term == "importerGermany" ~ "Germany",
    term == "importerJapan" ~ "Japan",
    term == "importerSouth Korea" ~ "South Korea",
    term == "importerUnited Kingdom" ~ "UK",
    term == "post2018:importerAustralia" ~ "Australia × Post-2018",
    term == "post2018:importerFrance" ~ "France × Post-2018",
    term == "post2018:importerGermany" ~ "Germany × Post-2018",
    term == "post2018:importerJapan" ~ "Japan × Post-2018",
    term == "post2018:importerSouth Korea" ~ "South Korea × Post-2018",
    term == "post2018:importerUnited Kingdom" ~ "UK × Post-2018",
    term == "(Intercept)" ~ "Constant",
    TRUE ~ term
  )) %>%
  mutate(
    stars = case_when(
      p.value < 0.001 ~ "***",
      p.value < 0.01 ~ "**",
      p.value < 0.05 ~ "*",
      p.value < 0.1 ~ ".",
      TRUE ~ ""
    ),
    Estimate = paste0(round(estimate, 3), stars),
    `Std. Error` = round(std.error, 3),
    `p-value` = round(p.value, 3)
  ) %>%
  select(Term = term, Estimate, `Std. Error`, `p-value`)

# Add model stats at the bottom
table_stats <- tibble(
  Term = c("Observations", "R-squared", "Adjusted R-squared"),
  Estimate = as.character(c(nobs(model_agg),
                            round(summary(model_agg)$r.squared, 3),
                            round(summary(model_agg)$adj.r.squared, 3))),
  `Std. Error` = NA, `p-value` = NA
)

# Combine and display
bind_rows(table_df, table_stats) %>%
  kable(format = "html", caption = "<span style='font-size:22px; font-weight:700; color:#003049;'>Strategic Imports from China</span>") %>%
  kable_styling(font_size = 13, full_width = FALSE, position = "left")
```

</div>
<!-- RIGHT: Interpretation -->
<div style="flex: 1.8; font-size: 26px; color: #003049; padding-top: 10px; line-height: 1.5;">

	•	🇺🇸 U.S. Post-2018: −0.268
	<br>
	→ U.S. reduced strategic imports by 23.5%
	<br>
	<br>
	•	🇰🇷 South Korea × Post-2018: +0.940
	<br>
	→ Diverged by 156% vs. U.S.
	<br>
	<br>
	•	🇩🇪 Germany × Post-2018: +0.614
	<br>
	→ Diverged by 84.8%
	<br>
	<br>
	•	🇬🇧 UK × Post-2018: +0.598
	<br>
	→ Diverged by 81.9%
	<br>
	<br>
	<em>Relative divergence, not absolute import growth<em>

</div>
</div>

---
# Visualization: Country Divergence

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=7.5, fig.height=3.2, out.width='100%', out.extra='style="margin-bottom:0;"'}
model_coef <- tidy(model_agg) %>%
  filter(grepl("post2018", term)) %>%
  mutate(
    country = gsub("post2018:importer", "", term),
    pct_change = round(100 * (exp(estimate) - 1), 1)
  )

model_coef_plot <- model_coef %>% filter(term != "post2018")

ggplot(model_coef_plot, aes(x = reorder(country, pct_change), y = pct_change)) +
  geom_col(fill = "tomato") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray40") +
  labs(
    title = "Change in Strategic Imports Post-2018 vs. U.S.",
    subtitle = "% Change Relative to U.S. (Aggregated Country-Year Model)",
    x = "Country", y = "% Difference from U.S."
  ) +
  coord_flip() +
  theme_minimal()
```
→Germany, South Korea, and the UK show the greatest divergence from U.S. decoupling 
---
# Robustness Check

### *I test whether results hold under:*
- Robust (heteroskedasticity-consistent) standard errors

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(lmtest)
library(sandwich)
library(broom)
library(dplyr)
library(knitr)
library(kableExtra)

# Robust SEs for interaction model
robust_se <- vcovHC(model_agg, type = "HC1")
robust_results <- coeftest(model_agg, vcov = robust_se)

# Convert to tidy format
robust_df <- tidy(robust_results) %>%
  filter(grepl("post2018", term)) %>%
  mutate(
    term = recode(term,
      "post2018" = "USA (Post-2018)",
      "post2018:importerAustralia" = "Australia × Post-2018",
      "post2018:importerFrance" = "France × Post-2018",
      "post2018:importerGermany" = "Germany × Post-2018",
      "post2018:importerJapan" = "Japan × Post-2018",
      "post2018:importerSouth Korea" = "South Korea × Post-2018",
      "post2018:importerUnited Kingdom" = "UK × Post-2018"
    ),
    stars = case_when(
      p.value < 0.001 ~ "&#9733;&#9733;&#9733;",   # ★★★
      p.value < 0.01 ~ "&#9733;&#9733;",           # ★★
      p.value < 0.05 ~ "&#9733;",                  # ★
      p.value < 0.1 ~ ".", 
      TRUE ~ ""
    ),
    Estimate = as.character(paste0(round(estimate, 3), stars)),
    `Std. Error` = round(std.error, 3),
    `p-value` = round(p.value, 3)
  ) %>%
  select(Term = term, Estimate, `Std. Error`, `p-value`)

# Render table
kable(robust_df,
      format = "html",
      escape = FALSE,
      caption = "Robust SEs (HC1): Post-2018 Effects by Country") %>%
  kable_styling(full_width = FALSE, font_size = 13, position = "left")
```
✅ Results are consistent across robustness checks.  
The divergence in allied behavior holds — even with alternative specifications.
---
# Key Takeaways

- U.S. strategic imports from China **declined** after 2018
- Most allies *did not* follow - several increased strategic imports
- Alliance cohesion appears limited in the economic domain

---
class: small-title
# Appendix: 13 Strategic Products (HS6 Codes)

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(knitr)
library(kableExtra)

# Create product table
strategic_products <- data.frame(
  Sector = c("Semiconductors", "", "", "", 
             "Telecom Equipment", "", 
             "Aerospace", "", 
             "Engine Components", 
             "Electronics", "", 
             "Batteries", 
             "Rare Earths"),
  HS_Code = c("854231", "854232", "854233", "854239",
              "851770", "852990",
              "880330", "880390",
              "840999",
              "850440", "847150",
              "850760",
              "280530"),
  Description = c("Microprocessor units", 
                  "Memory chips", 
                  "Amplifier ICs", 
                  "Other integrated circuits", 
                  "Telephone or networking parts", 
                  "Radio/telecom transmission parts", 
                  "Aircraft parts", 
                  "Other aircraft or helicopter parts", 
                  "Parts for aircraft or vehicle engines", 
                  "Power supply units and converters", 
                  "Computer processing units", 
                  "Lithium-ion batteries", 
                  "Yttrium oxide (rare earth metal)")
)

# Print table with smaller font
kable(strategic_products, "html", align = "lll") %>%
  kable_styling(font_size = 16, full_width = FALSE, position = "center")
```
*These HS6 codes were selected based on their prominence in U.S. trade policy debates, export controls, and strategic technology initiatives.

---

class: center, middle, question-slide

# Thank You!

## Questions?



*il0974a@american.edu*
<br>
*GitHub: ikhyun813*

