---
title: "Did U.S. Allies Follow Washington's Economic Decoupling?"
author: "[Ikhyun Lee]"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: [default, rladies, rutgers-fonts]
    nature:
      highlightStyle: github
      highlightLines: true
      ratio: "16:9"
---
# Overview

- **Motivation**: The U.S. launched a trade war with China in 2018, aiming to decouple economically, especially in strategic sectors (e.g., semiconductors, telecom).
- **Research Question**: Did U.S. allies follow suit by reducing their own strategic imports from China?
- **Finding**: Most U.S. allies *did not* follow the U.S.'s lead — many increased imports instead.

---
# Data

- **Source**: UN Comtrade (cleaned in R)
- **Years**: 2014–2022
- **Countries**: United States, Japan, South Korea, UK, Germany, France, Australia
- **Products**: 13 HS codes classified as strategic (e.g., semiconductors, aircraft parts, telecom equipment)
- **Unit of Analysis**: Country–year–product imports from China (USD)

---
# Analysis 1: Total Strategic Imports from China

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=5, out.width='100%'}
library(ggplot2)
library(scales)
library(dplyr)
library(stargazer)

df <- read.csv("cleaned_trade_data.csv")
df_summary <- df %>%
  group_by(year, importer) %>%
  summarise(total_import = sum(trade_value, na.rm = TRUE), .groups = "drop")

# Plot total imports by year-country
ggplot(df_summary, aes(x = year, y = total_import / 1e9, color = importer)) +
  geom_line(size = 1) +
  geom_point() +
  geom_vline(xintercept = 2018, linetype = "dashed", color = "red") +
  scale_y_continuous(labels = label_number(suffix = "B")) +
  labs(
    title = "Total Strategic Goods Imports from China",
    subtitle = "2014–2022 (Dashed Line: U.S.-China Trade War Begins)",
    x = "Year", y = "Import Value (USD, Billions)", color = "Country"
  ) +
  theme_minimal()
```

---
# Analysis 2: Post-2018 Divergence (Aggregated Interaction Model)

This model estimates how each country's strategic imports from China changed relative to the U.S. after 2018.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(broom)
library(knitr)
library(kableExtra)

# Use aggregated country-year data
df_agg <- df %>%
  group_by(importer, year) %>%
  summarise(
    trade_value = sum(trade_value, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  filter(trade_value > 0) %>%
  mutate(
    log_trade = log(trade_value),
    post2018 = ifelse(year >= 2018, 1, 0),
    importer = relevel(factor(importer), ref = "United States")
  )

model_agg <- lm(log_trade ~ post2018 * importer, data = df_agg)
```
---
# Regression Table
```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}

# Tidy + add significance stars
table_df <- tidy(model_agg) %>%
  mutate(term = case_when(
    term == "post2018" ~ "USA (Post-2018)",
    term == "importerAustralia" ~ "Australia",
    term == "importerFrance" ~ "France",
    term == "importerGermany" ~ "Germany",
    term == "importerJapan" ~ "Japan",
    term == "importerSouth Korea" ~ "South Korea",
    term == "importerUnited Kingdom" ~ "UK",
    term == "post2018:importerAustralia" ~ "Australia × Post-2018",
    term == "post2018:importerFrance" ~ "France × Post-2018",
    term == "post2018:importerGermany" ~ "Germany × Post-2018",
    term == "post2018:importerJapan" ~ "Japan × Post-2018",
    term == "post2018:importerSouth Korea" ~ "South Korea × Post-2018",
    term == "post2018:importerUnited Kingdom" ~ "UK × Post-2018",
    term == "(Intercept)" ~ "Constant",
    TRUE ~ term
  )) %>%
  mutate(
    stars = case_when(
      p.value < 0.001 ~ "***",
      p.value < 0.01 ~ "**",
      p.value < 0.05 ~ "*",
      p.value < 0.1 ~ ".",
      TRUE ~ ""
    ),
    Estimate = paste0(round(estimate, 3), stars),
    `Std. Error` = round(std.error, 3),
    `p-value` = round(p.value, 3)
  ) %>%
  select(Term = term, Estimate, `Std. Error`, `p-value`)

# Add model stats at the bottom
table_stats <- tibble(
  Term = c("Observations", "R-squared", "Adjusted R-squared"),
  Estimate = as.character(c(nobs(model_agg),
                            round(summary(model_agg)$r.squared, 3),
                            round(summary(model_agg)$adj.r.squared, 3))),
  `Std. Error` = NA, `p-value` = NA
)

# Combine and display
bind_rows(table_df, table_stats) %>%
  kable(format = "html", caption = "Regression Results: Strategic Imports from China") %>%
  kable_styling(font_size = 12, full_width = FALSE)
```

---
# Visualization: Country Divergence

```{r, echo=FALSE, message=FALSE, fig.width=10, fig.height=5, out.width='100%'}
model_coef <- tidy(model_agg) %>%
  filter(grepl("post2018", term)) %>%
  mutate(
    country = gsub("post2018:importer", "", term),
    pct_change = round(100 * (exp(estimate) - 1), 1)
  )

model_coef_plot <- model_coef %>% filter(term != "post2018")

ggplot(model_coef_plot, aes(x = reorder(country, pct_change), y = pct_change)) +
  geom_col(fill = "tomato") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray40") +
  labs(
    title = "Change in Strategic Imports Post-2018 vs. U.S.",
    subtitle = "% Change Relative to U.S. (Aggregated Country-Year Model)",
    x = "Country", y = "% Difference from U.S."
  ) +
  coord_flip() +
  theme_minimal()
```

---
# Takeaways

- **U.S. imports declined** after 2018 (though not always significantly)
- **Germany, Korea, UK** showed clear divergence with increased imports
- **Conclusion**: Despite U.S. decoupling efforts, allies' behaviors were not cohesive

---
# Thank You!

Questions?

**Email**: [il0974a@american.edu]  
**GitHub**: [ikhyun813]
